Learning Augmented Joint-Space Task-Oriented Dynamical Systems: \\ A Linear Parameter Varying and Synergetic Control Approach

### 2018-03-11 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
changes in paper

\begin{abstract}
In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while stably converging towards a task-space attractor. % Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor. 
 To encode joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations. Specifically, we use dimensionality reduction to find a low-dimensional embedding space for modulating joint synergies, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations. Our proposed approach is empirically validated on a variety of motions. %for a 7-DOF KUKA LWR4+ robot arm.
 \end{abstract}

@@

Many LfD approaches learn motions in either joint space %\cite{Calinon:HRI:2007, Garrido:Neuro:2015, Schaal:IROS:2003,Schaal:AI:2002}
\cite{Garrido:Neuro:2015,Schaal:AI:2002} or task space %\cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}
\cite{Pastor:ICRA:2009,Calinon:ISR:2015}. These approaches generally learn probabilistic model of the demonstrated motions. %, and construct a dynamical system that chooses behaviors according to the learned model.
Both %joint and task 
spaces are independently useful:  task-space behaviors let us control the most task-relevant component of the robot's behavior. In particular, a number of approaches have pioneered stable convergence to an attractor in task space % \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015} 
 \cite{khansari2011learning, Neumann:RAS:2015}as a desirable capability for a dynamical system.

@@

Furthermore, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations. The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}. When the main focus is on executing a non-linear task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}. However, for other applications, such an approach yields significant problems \cite{buss2004introduction}. First of all, finding the pseudo-inverse is computationally taxing. Moreover, when the Jacobian matrix cannot be inverted (i.e. when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior. These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}. By defining a controller in joint space, one can avoid IK and all its associated drawbacks.

@@

%\subsection{Our Proposed Approach}
In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations. To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system. We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories. 
Hence, our dynamical system:
\begin{enumerate}[leftmargin=*]
\item computes a \begin{bf}motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
\item is formulated such that \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
\item can transit through kinematic \textbf{singularities}.
\end{enumerate}

@@

%\begin{figure}[t]
%\centering
%\includegraphics[width=0.95\linewidth,trim={0 20 450 0},clip]{Pic/Schematic.jpg}
%\includegraphics[width=0.9\linewidth,trim={580 20 0 0},clip]{Pic/Schematic.jpg}
%\caption{ Standard control architecture using Cartesian dynamical systems (DS)\textbf{(top)}  and our proposed approach \textbf{(bottom)}.}
%\label{fig:schematic}
%\end{figure}
\section{Problem Statement} \label{Sec:Prob}
Consider a robotic system with $d$ task-space dimensions and $m$ degrees of freedom. We direct the system using a joint position or joint velocity controller, which can have joint position limits and a maximum joint velocity. We are further provided with a set of $N$ demonstrated joint-space trajectories $D=\{\{q_{t,n},\dot{q}_{t,n}\}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$, where $T_n$ is the number of the sample points of the $n^{\text{th}}$ demonstration. We refer to the system's joint-space position as $q=\begin{bmatrix} q^1 & \dots & q^m
\end{bmatrix}^T \in \mathbb{R}^m$, and to its task-space position as $x\in \mathbb{R}^d$.\footnote{For sake of brevity and simplicity, the time index, $t$, is dropped throughout the paper. } The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{m\times d}$.\\
We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following two criteria:
\renewcommand{\labelenumi}{(\Roman{enumi})}
\begin{enumerate}
\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified,``stability" in this paper always refers to asymptotic stability within the workspace of the robot, and assumes no joint limits. We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system.} with respect to a fixed task-space target $x^*$. This can be expressed by ensuring that the following Lyapunov function
\begin{equation}
V(q) = (H(q) - x^*)^T(H(q) - x^*) 
\label{eq:Lyp}
\end{equation}
is stable; i.e. $\dot{V}(q)\leq 0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$. Where  $Q=\{q|q^i_{min} < q^i < q^i_{max},~\forall i\in \{1,\dots,d\}\}$ and $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$. $P\in \mathbb{R}^{d\times d} $ is a symmetric and positive definite matrix. $V(.)$ can be thought of as a metric for the task-space distance-to-go.
\item The dynamical system should encapsulate the desired joint-space behaviors such that the following metric is minimized
\begin{equation}
\vspace{-5pt}
e_{total} = \frac{1}{NT_n}\sum_{n=1}^N\sum_{t=0}^{T_n} \left \| \dot{q}_{d; t,n} - f(q_{t,n}) \right \|
\label{eq:optim_error}
\end{equation}
  where $\dot{q}_d$ is the desired ``true" velocity, and $f(\cdot)$ is the motion generation policy.
\end{enumerate}

@@

The error metric \eqref{eq:optim_error} is advantageous in a way that it mimics the magnitude and direction of the demonstrated motions. Hence, as suggested by \cite{gielniak2010stylized}, it is visually most similar to the human definition of ``joint motion style".

@@

%We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}. First, the metric implicitly suggests that behaviors are defined by expert motions of the behavior, and that fulfilling a behavior means moving similarly to the experts' motion. Second, the type of feature of the trajectory that is being minimized is important. If instead the error feature were joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity. Alternatively, if the error feature were joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior). By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
\section{Augmented \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} \label{Sec:DS}
\label{sec:proposed_system}
We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \eqref{eq:Lyp} and \eqref{eq:optim_error}.
\begin{equation}
\label{eq:ds}
\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)(H(q) - x^*)
\end{equation}
where  $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the LPV system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy. 
\begin{equation}
\label{eq:A_def}
\mathcal{A}(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
\end{equation}
where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{%The scheduling parameters can be a function of time $t$, the states of the
%system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. 
It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$, where $d(t)$ is an external signal.} modulating each local synergy through time as well as space. Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more non-linear multi-behavior motion.

Before proving that the proposed dynamical system satisfies all of the criteria, let us establish an intuitive understanding of the components of the system. While it may seem daunting at first, each of the elements has a straightforward explanation. One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig. \ref{fig:error_schematic}. The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig. \ref{fig:A_example} illustrates the effects of  $\mathcal{A}(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space. Lastly, we refer to %the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(as it augments the task error, and thus the direction of motion) and 
$\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} as it augments the outputted joint velocities.
%When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle). This ''augmentation" is simply a matrix multiplication ($A\dot{q}_{raw}$). The components of the outputted velocity are linear combinations of the ''raw" un-augmented joint velocities, where the entries in $A$ define the precise ratios. We can see the joint positions resulting from our experiment over time on the right side of Fig. \ref{fig:A_example}. Each motion tends to use its ``primary" joint most (the joint where the velocity is heavily scaled up) and uses the other available joints to compensate for what the primary joint cannot do. For example, Motion 1 starts out using mostly the hip, but once moving the hip no longer brings the system closer to the goal, the system uses the other joints to fold inwards and reach the goal. Ultimately, all three controllers converge on their target, but each has a different path through joint space depending on which joints we choose to emphasize in the behavior matrix $A$. The off-diagonal terms of $A$ have a more nuanced contribution; e.g. $A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.

\newtheorem{prop1}{Proposition}
\begin{prop1}
\label{prop:stability}
The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ \eqref{eq:ds} meets the following constraints. 
\begin{equation}
\label{eq:first_criteria_stability}
\begin{cases}
\begin{split}
 & 0 \prec A_k, & ~~~~ 0 \leq \theta_k(\cdot) \\
\end{split}
\end{cases}
\end{equation}    
\end{prop1}
\vspace{5pt}
\textbf{Proof}: See Appendix \ref{appendix:stability}. $\blacksquare$
%It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  only ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$ if the robot does not have kinematic joint limits. If we introduce kinematic joint limits, then it's possible we may find $\exists q\in Q-Q^* $ such that both $\dot{V}(q)=0$ and $H(q) \neq x^*$. This happens if the target position is kinematically unreachable; i.e. for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or the joint has reached a kinematic limit. In this case, the proof in Appendix \ref{appendix:stability} shows instead that the system converges in the sense of Lyapunov, meaning that its distance to the attractor is always monotonically decreasing.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
	\caption{Illustration of our proposed Lyapunov function and error projection. On the left we show the task-space potential function that we are trying to descend on for a 2-DOF robot with target $x^*=[2,2]^T$. On the right we show the corresponding Lyapunuv function in joint space \eqref{eq:Lyp}. The colors on each plot correspond to each other, one can see how the error, and consequently the \textit{attractive} regions, in task-space are warped in joint-space.}
	\label{fig:error_schematic}
	\vspace{-15pt}
\end{figure}

@@

In this work, we assume that configurations that are nearby in joint-space should exhibit similar behaviors, and thus it is natural that we choose a low-dimensional embedding that preserves the variance in our demonstrations. To this end, we construct an embedding $\phi(q)$ by training dimensionality reduction techniques on the demonstrated trajectories. The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^d$ into a lower-dimensional configuration $z \in \mathbb{R}^\delta$, where $\delta < d$.
%Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}. 
%In previous work, PCA has shown to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}, and this work further explores the usefulness of joint-space dimensionality reduction.

@@

Scheduling parameters for LPV systems with such properties have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}. Intuitively, we search for a probabilistic model that ``explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a local behavior, which the synergy will then approximate. In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the scheduling parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and covariance matrices that parametrize the $k$-th multivariate Gaussian distribution. Each distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ to the $k$th synergy of \eqref{eq:A_def}. We define $\theta'_k(\phi(q))$ as $p(k|\phi(q))$:
\begin{equation}
\label{eq:theta}
\theta'_k(\phi(q))= \frac{\pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}
\end{equation}
which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th synergy. Therefore, each synergy region is associated with a Gaussian component of the GMM, cumulatively describing all the synergy regions of the dynamical system. We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM.

@@

Once these optimal hyper-parameters are estimated for each fold, we solve for the convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}. We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}

>>>>

\begin{abstract}
In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a task-space attractor. % Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor. 
 To encode joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations. Specifically, we use dimensionality reduction to find a low-dimensional embedding space for modulating joint synergies, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations. Our proposed approach is empirically validated on a variety of motions. %for a 7-DOF KUKA LWR4+ robot arm.
 \end{abstract}

@@

Many LfD approaches learn motions in either joint space %\cite{Calinon:HRI:2007, Garrido:Neuro:2015, Schaal:IROS:2003,Schaal:AI:2002}
\cite{Garrido:Neuro:2015,Schaal:AI:2002} or task space %\cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}
\cite{Pastor:ICRA:2009,Calinon:ISR:2015}. These approaches generally learn a probabilistic model of the demonstrated motions. %, and construct a dynamical system that chooses behaviors according to the learned model.
%Both joint and task spaces are independently useful:  task-space behaviors let us control the most task-relevant component of the robot's behavior. 
A number of approaches have further proposed stable convergence to an attractor in task space % \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015} 
 \cite{khansari2011learning, Neumann:RAS:2015}as a desirable capability for a dynamical system.

@@

Furthermore, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations. The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}. When the main focus is on executing a non-linear task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}. However, for other applications, such an approach yields significant problems \cite{buss2004introduction}. First of all, finding the pseudo-inverse is computationally taxing \textbf{insert how slow it is for a 1000Hz robot}. Moreover, when the Jacobian matrix cannot be inverted (i.e. when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior. These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}. By defining a controller in joint space, one can avoid IK and all its associated drawbacks.

@@

%\subsection{Our Proposed Approach}
In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations. To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system. We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories. 
Hence, our dynamical system:
\begin{enumerate}[leftmargin=*]
\item computes a \begin{bf}motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
\item is formulated such that \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
\item can transit through kinematic \textbf{singularities}.
\end{enumerate}

@@

%\begin{figure}[t]
%\centering
%\includegraphics[width=0.95\linewidth,trim={0 20 450 0},clip]{Pic/Schematic.jpg}
%\includegraphics[width=0.9\linewidth,trim={580 20 0 0},clip]{Pic/Schematic.jpg}
%\caption{ Standard control architecture using Cartesian dynamical systems (DS)\textbf{(top)}  and our proposed approach \textbf{(bottom)}.}
%\label{fig:schematic}
%\end{figure}
\section{Problem Statement} \label{Sec:Prob}
Consider a robotic system with $d$ task-space dimensions and $m$ degrees of freedom. We direct the system using a joint position or joint velocity controller, which can have joint position limits and a maximum joint velocity. We are further provided with a set of $N$ demonstrated joint-space trajectories $D=\{\{q_{t,n},\dot{q}_{t,n}\}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$, where $T_n$ is the number of the sample points of the $n^{\text{th}}$ demonstration. We refer to the system's joint-space position as $q=\begin{bmatrix} q^1 & \dots & q^m
\end{bmatrix}^T \in \mathbb{R}^m$, and to its task-space position as $x\in \mathbb{R}^d$.\footnote{For sake of brevity and simplicity, the time index, $t$, is dropped throughout the paper. } The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{m\times d}$.\\
We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following two criteria:
\renewcommand{\labelenumi}{(\Roman{enumi})}
\begin{enumerate}
\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified, ``stability" in this paper always refers to asymptotic stability within the workspace of the robot (in the regions where the Jacobian is full rank), and as such assumes no joint limits. We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system.} with respect to a fixed task-space target $x^*$. This can be expressed by ensuring that the following Lyapunov function
\begin{equation}
V(q) = (H(q) - x^*)^T(H(q) - x^*) 
\label{eq:Lyp}
\end{equation}
is stable; i.e. $\dot{V}(q) < 0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$ where $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$ and $Q=\{q|q\notin Q^*, J(q) \textit{ is full rank}\}$. $V(.)$ can be thought of as a metric for the task-space distance-to-go.
\item The dynamical system should encapsulate the desired joint-space behaviors such that the following metric is minimized
\begin{equation}
\vspace{-5pt}
e_{total} = \frac{1}{NT_n}\sum_{n=1}^N\sum_{t=0}^{T_n} \left \| \dot{q}_{d; t,n} - f(q_{t,n}) \right \|
\label{eq:optim_error}
\end{equation}
  where $\dot{q}_d$ are the ``true" velocities from the demonstrations, and $f(\cdot)$ is the motion generation policy.
\end{enumerate}

@@

The error metric \eqref{eq:optim_error} is advantageous in that it mimics the magnitude and direction of the demonstrated motions. Hence, as suggested by \cite{gielniak2010stylized}, it is visually most similar to the human definition of ``joint motion style".

@@

%We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}. First, the metric implicitly suggests that behaviors are defined by expert motions of the behavior, and that fulfilling a behavior means moving similarly to the experts' motion. Second, the type of feature of the trajectory that is being minimized is important. If instead the error feature were joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity. Alternatively, if the error feature were joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior). By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
\section{Augmented \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} \label{Sec:DS}
\label{sec:proposed_system}
We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \eqref{eq:Lyp} and \eqref{eq:optim_error}.
\begin{equation}
\label{eq:ds}
\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)(H(q) - x^*)
\end{equation}
where  $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the LPV system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy. Each ``synergy" is a joint-space transformation that biases the resulting motion to use particular joints; for more, see Sec. \ref{Sec:Learning}. Explicitly,
\begin{equation}
\label{eq:A_def}
\mathcal{A}(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
\end{equation}
where $A_k\in \mathbb{R}^{m\times m} $ are the different synergy matrices and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{%The scheduling parameters can be a function of time $t$, the states of the
%system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. 
It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$, where $d(t)$ is an external signal.} modulating each local synergy through time as well as space. Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more non-linear multi-behavior motion.

Before proving that the proposed dynamical system satisfies all of the criteria, let us establish an intuitive understanding of the components of the system. While it may seem daunting at first, each of the elements has a straightforward explanation. One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target, which we then reinterpret as the task-space velocity of a proportional controller. By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig. \ref{fig:error_schematic}. The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig. \ref{fig:A_example} illustrates the effects of  $\mathcal{A}(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space. Lastly, we refer to %the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(as it augments the task error, and thus the direction of motion) and 
$\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} as it augments the outputted joint velocities.
%When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle). This ''augmentation" is simply a matrix multiplication ($A\dot{q}_{raw}$). The components of the outputted velocity are linear combinations of the ''raw" un-augmented joint velocities, where the entries in $A$ define the precise ratios. We can see the joint positions resulting from our experiment over time on the right side of Fig. \ref{fig:A_example}. Each motion tends to use its ``primary" joint most (the joint where the velocity is heavily scaled up) and uses the other available joints to compensate for what the primary joint cannot do. For example, Motion 1 starts out using mostly the hip, but once moving the hip no longer brings the system closer to the goal, the system uses the other joints to fold inwards and reach the goal. Ultimately, all three controllers converge on their target, but each has a different path through joint space depending on which joints we choose to emphasize in the behavior matrix $A$. The off-diagonal terms of $A$ have a more nuanced contribution; e.g. $A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.

\newtheorem{prop1}{Proposition}
\begin{prop1}
\label{prop:stability}
The motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ \eqref{eq:ds} meets the following constraints. 
\begin{equation}
\label{eq:first_criteria_stability}
\begin{cases}
\begin{split}
 & A_k \succ 0, & ~~~~ \theta_k(\cdot) \geq 0 \\
\end{split}
\end{cases}
\end{equation}    
\end{prop1}
\vspace{5pt}
\textbf{Proof}: See Appendix \ref{appendix:stability}. $\blacksquare$
%It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  only ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$ if the robot does not have kinematic joint limits. If we introduce kinematic joint limits, then it's possible we may find $\exists q\in Q-Q^* $ such that both $\dot{V}(q)=0$ and $H(q) \neq x^*$. This happens if the target position is kinematically unreachable; i.e. for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or the joint has reached a kinematic limit. In this case, the proof in Appendix \ref{appendix:stability} shows instead that the system converges in the sense of Lyapunov, meaning that its distance to the attractor is always monotonically decreasing.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
	\caption{Illustration of our task-space potential function and proposed joint-space Lyapunov function. On the left we show the task-space error we are trying to minimize, represented by a potential function for a 2-DOF robot with target $x^*=[2,2]^T$. On the right we show the corresponding Lyapunov function in joint space defined in \eqref{eq:Lyp}. The colors on each plot correspond to each other. One can see how the error, and consequently the \textit{attractive} regions in task-space, have been warped into attractive regions in joint-space.}
	\label{fig:error_schematic}
	\vspace{-15pt}
\end{figure}

@@

In this work, we assume that configurations that are nearby in joint-space should exhibit similar behaviors, and thus it is natural that we choose a low-dimensional embedding that preserves the variance in our demonstrations. To this end, we construct a single global embedding $\phi(q)$ by training dimensionality reduction techniques on the demonstrated trajectories. The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^d$ into a lower-dimensional configuration $z \in \mathbb{R}^\delta$, where $\delta < d$. For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the ``shoulder" and ``arm" components of $q$ into a single ``shoulder-arm" component in $\phi(q)$.
%Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}. 
%In previous work, PCA has shown to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}, and this work further explores the usefulness of joint-space dimensionality reduction.

@@

Scheduling parameters for LPV systems with such properties have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}. Intuitively, we search for a probabilistic model that ``explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a local behavior, which the synergy will then approximate. In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the scheduling parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and covariance matrices that parametrize the $k$-th multivariate Gaussian distribution. Each distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ to the $k$th synergy of \eqref{eq:A_def}. We define $\theta'_k(\phi(q))$ as $p(k|\phi(q))$:
\begin{equation}
\label{eq:theta}
\theta'_k(\phi(q))= \frac{\pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}
\end{equation}
which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th synergy. Therefore, each synergy region is associated with a Gaussian component of the GMM, cumulatively describing all the synergy regions of the dynamical system. We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM \textbf{ADD CITATION FOR EM}.

@@

Once these optimal hyper-parameters are estimated for each fold, we solve for the convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}. We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}.
